\chapter{PLGraphics}


\paragraph{Motivation}
There are tons of image file formats and image libraries. Images are a fundamental feature of GUI and 3D rendering systems. Because they are that important, we decided to create \emph{PLGraphics} which is an image library that fulfils the requirements of PixelLight and integrates itself quite well into the whole framework. Beside the image features, other important feature are the color classes.




\section{External Dependences}
The core of PLGraphics depends on the \textbf{PLCore} and \textbf{PLMath} libraries. PLGraphics comes with some static linked image loader plugins so the library can directly load in several standard format. Some image loader plugins are using open source third party libraries which are static linked. As a result, the resulting PLGraphics library has no additional external third party shared library dependencies.


\paragraph{libjpeg}
\begin{itemize}
\item JPG image loading library
\item \emph{jpg} license
\item Version \emph{8b} (16-May-2010)
\item Used by the image loader plugin class \emph{ImageLoaderJPG}
\item Downloaded from \url{http://www.ijg.org/}
\end{itemize}


\paragraph{libpng}
\begin{itemize}
\item PNG image loading library
\item \emph{png} license
\item Version \emph{1.4.8} (July 7, 2011)
\item Used by the image loader plugin class \emph{ImageLoaderPNG}
\item Downloaded from \url{http://www.libpng.org/}
\end{itemize}




% Include the other sections of the chapter
\input{PLGraphics/Image}
\cleardoublepage
