\chapter{Misc}




\section{Toolbar}
By using \emph{Utilities} -> \emph{More...} -> \emph{PixelLight Scene Exporter Toolbar} you can add a toolbar into \emph{Autodesk 3ds Max} to make PixelLight previews quite comfortable.




\section{Tips}
\begin{itemize}
\item{The \emph{center} of an object should be the real center of the object\footnote{Excep there's a good reason for another setting} - this is no must, but this can get quite uncomfortable if the center of an object is somewhere totally outside of the object. Within \emph{Autodesk 3ds Max}, select a node, select the hierarchy tab, within \emph{Adjust Pivot} select \emph{Affect Pivot Only}, within \emph{Alignment} select \emph{Center to Object}. Now the center of the node should be fine.}
\item{All object names should be unique - else the export of such objects with the same name may fail because the object name is also used to identify the used mesh. If there are objects with the same names within your scene, a warning is written into the log.\footnote{There's no different between upper/lower case letters within \emph{Autodesk 3ds Max}, but within PixelLight}}
\item{You should take care of upper/lower case letters. For instance if you use the texture \emph{MyTexture.dds}, don't load a texture named \emph{mYtExTUrE.DDS} - this would help no one and in case that something is case sensitive problems may occur!}
\item{Give your objects PRACTICAL names - this will make it more comfortable to edit the scenes using a text editor or a scene editor... a scene with objects named \emph{x0}-\emph{x10000} would be horror!}
\item{If the scene is rendered by using per pixel lighting, normally, correct per vertex normals, tangents, binormals and texture coordinates are required. If those per vertex normals and texture coordinates are not correct, the resulting lighting within PixelLight may look not correct.}
\item{Try to avoid using none uniform scaling for nodes because they may cause problems in special situations... for instance if a sphere radius should be multiplied with the node scale and should not result in an ellipsoid. If a none uniform scale is found, a \emph{hint} is written into the log.}
\item{Try to avoid using negative scaling for nodes because it may cause problems.}
\item{When using the \emph{Basic.sr} scene renderer, ensure that \emph{Ambient Color} within \emph{Autodesk 3ds Max} is not black, because this scene renderer has no support for lights and as result the lighting is heavy depending on this \emph{Ambient Color}.}
\end{itemize}
