\section{Sound Backends}
This section deals with the different sound backends shipping with the PixelLight SDK.


\paragraph{Null}
\begin{itemize}
\item The null sound backend does nothing - can be useful if you e.g. are in debug mode and don't want to hear any sounds.
\item PixelLight component name: PLSoundNull
\item Used dll's: \emph{PLSoundNull.dll}
\end{itemize}


\paragraph{OpenAL}
\begin{itemize}
\item OpenAL\footnote{OpenAL can be downloaded from \url{http://www.openal.org/}} backend with support of wav and ogg vorbis (streaming/no streaming) is provided.
\item PixelLight component name: PLSoundOpenAL
\item Used dll's: \emph{PLSoundOpenAL.dll}, \emph{OpenAL32.dll} and \emph{wrap\_oal.dll}
\end{itemize}


\paragraph{FMODEx}
\begin{itemize}
\item Currently not within the official PixelLight SDK
\item FMODEx\footnote{FMODEx can be downloaded from \url{http://www.fmod.org/}, only free for none commercial projects, \emph{FMOD Sound System, copyright © Firelight Technologies Pty, Ltd., 1994-2011.}} backend which can produce a lot of state of the art sound effects! Some of many supported formats are mp3, wav, mid, midi, it, mod, s3m, xm etc. 
\item PixelLight component name: PLSoundFMODEx
\item Used dll's: \emph{PLSoundFMODEx.dll} and \emph{fmodex.dll}
\end{itemize}


\paragraph{FMOD}
\begin{itemize}
\item Currently not within the official PixelLight SDK
\item FMOD\footnote{FMOD can be downloaded from \url{http://www.fmod.org/}, only free for none commercial projects, \emph{FMOD Sound System, copyright © Firelight Technologies Pty, Ltd., 1994-2011.}} backend which can produce a lot of state of the art sound effects! Some of many supported formats are mp3, wav, mid, midi, it, mod, s3m, xm etc. 
FMOD is the predecessor of FMODEx. Use the newer FMODEx if possible.
\item PixelLight component name: PLSoundFMOD
\item Used dll's: \emph{PLSoundFMOD.dll} and \emph{fmod.dll}
\end{itemize}
