\chapter{PLGui RTTI Classes}




\section{PLGui::Gui Class}


\subsection{Methods}

\paragraph{SetMouseVisible(<bool>)}
Write \emph{<RTTI object>:SetMouseVisible(<bool>)} in order to set the mouse cursor visibility. \emph{true} as first parameter if the mouse cursor shall be visible.




\section{PLGui::Widget Class}


\subsection{Attributes}

\paragraph{ID}
Widget ID.

\paragraph{Name}
Widget name.

\paragraph{Title}
Widget title.

\paragraph{Pos}
Widget position.

\paragraph{Size}
Widget size.

\paragraph{Topmost}
Topmost state.

\paragraph{BackgroundColor}
Background color.


\subsection{Methods}

\paragraph{<RTTI object> GetGui()}
Write \emph{<RTTI object>:GetGui()} in order to get the owner GUI. Returns pointer to GUI object, can be a null pointer.

\paragraph{<RTTI object> GetContentWidget()}
Write \emph{<RTTI object>:GetContentWidget()} in order to get the content widget, can be a null pointer.

\paragraph{SetTrapMouse(<bool>)}
Write \emph{<RTTI object>:SetTrapMouse(<bool>)} in order to trap the mouse inside the widget. \emph{true} as first parameter if the mouse should be trapped inside the widget, else \emph{false}.


\subsection{Signals}

\paragraph{SignalUserMessage}
User message, two custom parameters.

\paragraph{SignalOnThemeChanged}
Theme has been changed.

\paragraph{SignalUpdateContent}
Widget content has been changed.

\paragraph{SignalUpdateChildWidget}
Child widget has been changed, child widget as parameter.

\paragraph{SignalAddChildWidget}
Child widget has been added, child widget as parameter.

\paragraph{SignalRemoveChildWidget}
Child widget has been removed, child widget as parameter.

\paragraph{SignalClose}
Widget shall be closed (ALT+F4 or X-Button pressed).

\paragraph{SignalCreate}
Widget has just been created.

\paragraph{SignalDestroy}
Widget is going to be destroyed.

\paragraph{SignalShow}
Widget gets shown.

\paragraph{SignalHide}
Widget gets hidden.

\paragraph{SignalEnable}
Widget gets enabled.

\paragraph{SignalDisable}
Widget gets disabled.

\paragraph{SignalGetFocus}
Widget gets focus.

\paragraph{SignalLooseFocus}
Widget looses focus.

\paragraph{SignalActivate}
Widget has been activated or deactivated (focus-widget has changed), state as parameter.

\paragraph{SignalDrawBackground}
Widget background gets drawn, graphics object used for painting as parameter.

\paragraph{SignalDraw}
Widget gets drawn, graphics object used for painting as parameter.

\paragraph{SignalMove}
Widget gets moved, new widget position as parameter.

\paragraph{SignalSize}
Widget gets resized, new widget size as parameter.

\paragraph{SignalWindowState}
Window state has changed, new window state as parameter.

\paragraph{SignalEnterFullscreen}
Widget enters fullscreen mode.

\paragraph{SignalLeaveFullscreen}
Widget leaves fullscreen mode.

\paragraph{SignalPreferredSize}
Widget has calculated it's preferred size, preferred size as parameter.

\paragraph{SignalAdjustContent}
Widget content has to be adjusted.

\paragraph{SignalMouseEnter}
Mouse enters widget.

\paragraph{SignalMouseLeave}
Mouse leaves widget.

\paragraph{SignalMouseOver}
Mouse-over widget has changed, state as parameter.

\paragraph{SignalMouseMove}
Mouse moves inside the widget, mouse position within the widget as parameter.

\paragraph{SignalMouseHover}
Mouse hovers over the widget.

\paragraph{SignalMousePosUpdate}
Mouse position inside the widget has changed due to the movement of widget, position as parameter.

\paragraph{SignalMouseButtonDown}
Mouse button is pressed, mouse button and mouse position within the widget as parameters.

\paragraph{SignalMouseButtonUp}
Mouse button is released, mouse button and mouse position within the widget as parameters.

\paragraph{SignalMouseButtonClick}
Mouse button has been clicked, mouse button and mouse position within the widget as parameters.

\paragraph{SignalMouseButtonDoubleClick}
Mouse button has been double-clicked, mouse button and mouse position within the widget as parameters.

\paragraph{SignalMouseWheel}
Mouse wheel moved, mouse wheel movement as parameter.

\paragraph{SignalKeyDown}
Key gets pressed, pressed key and modifier keys pressed as parameters.

\paragraph{SignalKeyUp}
Key gets released, released key and modifier keys pressed as parameters.

\paragraph{SignalHotkey}
Hotkey pressed, hotkey ID as parameter.

\paragraph{SignalDrop}
Data has been dropped onto the widget, data as parameter.




\section{PLGui::GuiApplication Class}


\subsection{Methods}

\paragraph{<RTTI object> GetMainWindow()}
Write \emph{<RTTI object>:GetMainWindow()} in order to get the main window. Returns pointer to the main window of the application, a null pointer on error.

\paragraph{SetMainWindow(<RTTI object>)}
Write \emph{<RTTI object>:SetMainWindow(<RTTI object>)} in order to set the main window, pointer to the main window of the application (a null pointer is also valid) as first parameter.
