\chapter{Script}
Because Lua is shipped with the official PixelLight \ac{SDK}, this script language will be used within the examples. Please note that the PixelLight script \ac{API} is script language independent. In fact, there's not even such a PixelLight script \ac{API}. Everything that is connected to the \ac{RTTI} of PixelLight can be accessed through script languages as long as the script backend has support for \ac{RTTI} objects.

Within the PLCore documentation, there's a lot of inside detail information on how the \ac{RTTI} works. The script support itself is implemented within PLCore because scripting is heavily using PLCore features like the \ac{RTTI}. Therefore, adding script bindings or using \ac{RTTI} objects within scripts is fairly straightforward and doesn't require the writing of thousands of proxy/wrapper classes exposing C++ functionality to script languages.

Certain non-\ac{RTTI} parts of PixelLight are exposed to script languages through the loose plugin \emph{PLScriptBindings}. Please note that within PixelLight, scripting is completely optional, not mandatory - unlike some other engines out there were one is only able to use scripting. So, you can use scripting, but you are not forced to do so.

In general, the abstract script interface of PixelLight supports the following script features:
\begin{itemize}
\item{Global variables (with namespace support)}
\item{Global functions, C++ calls script and script calls C++ (with namespace support)}
\item{\ac{RTTI} objects}
\end{itemize}

As you will see when reading this document and playing around with the script support, the access to \ac{RTTI} objects makes the script support quite powerful and universal. You're able to access properties (constant \ac{RTTI} class information), attributes (variables within an object), methods (functions within an object), signals (events within an object) and slots (event handlers within an object).

Supported primitive data types are: \begin{quote}bool, float, double, int8, int16, int32, int64, uint8, uint16, uint32, uint64, PLCore::Object*, PLCore::Object\&\end{quote}

Please note that not each script language/\ac{API} may make such a detailed data type distinction. Because strings are fundamental within scripts, PLCore::String is supported as well. But if you're a 100\% script programmer, without writing own new C++ components, the mentioned information is probably already too technically.

In general, the script support of PixelLight can be subdivided into the two following use-cases:
\begin{itemize}
\item{A C++ component is using an own script instance in a highly specialized way. The script scene node modifier is a good example for this. Such a script scene node modifier is attached to a scene node and is adding logic through a script.}
\item{A scripted stand-alone application, meaning that the complete application logic is implemented within, for instance Lua, and can be executed by \emph{PLViewer} which comes with the PixelLight \ac{SDK} or custom applications. This doesn't mean that it's 100\% script only because the more complex stuff will probably done by used C++ components, but the wires are tied up by a script.}
\end{itemize}




\section{Namespaces}
\label{Script_Namespaces}
When a project is growing, one may run out of decent names or name conflicts happen by accident more often. Even if you're just starting a project and don't think that you'll end up with much source code, you should consider to use namespaces right from the beginning. That's the reason why this section comes even before introducing, for instance, global variables.

Most modern high-level computer languages support the concept of namespaces which, when used in the proper way, can help a developer with the naming task. An even bigger benefit in using namespaces than reducing name conflicts is, that elements are grouped together. An editor may, for example, just look for global variables inside a namespace called \emph{PublicVariables} in order to make them visible and editable through a \ac{GUI} - while other internal script variables will be hidden from the editor user. This way, it's not required to \emph{extend} already existing, maybe even normed script languages like JavaScript (ECMA-262), by own constructs like a \emph{public}-modifier to denote that a script variable should be accessible to the outside world. Please note that this was just an example, within the PixelLight core there's nothing like \emph{PublicVariables} fixed build in.

While, on the first look, several script languages have no namespace support, they have so on the second look or it can be at least emulated. Let's take Lua as an example. Lua is a really compact language and many important features are missing within the core of the language... but on a deeper look they can be realized by using Lua's powerful and universal table concept. This is also true for namespaces.

When searching the internet for information how to do namespaces in Lua, one often finds examples like seen in source~code~\ref{Code:GlobalVariablesWithinANamespace1}.
\begin{lstlisting}[language=lua,float=htb,label=Code:GlobalVariablesWithinANamespace1,caption={Global variables within a namespace (1)}]
-- Create the "PublicVariables"-namespace
PublicVariables = {}
-- Create the variable "Health" within
-- the "PublicVariables"-namespace
PublicVariables.Health	= 42
-- Create the variable "Gold" within
-- the "PublicVariables"-namespace
PublicVariables.Gold	= 21
\end{lstlisting}
By writing \emph{PublicVariables = \{\}}, a new global variable named \emph{PublicVariables} is created and an empty table \emph{\{\}} is assigned to it. The name \emph{PublicVariables} is now addressing a Lua table. By writing \emph{PublicVariables.Health = 42} the table \emph{PublicVariables} gets a new entry with the key \emph{Health} and the value \emph{42}.

Personally, I don't like the style used in source~code~\ref{Code:GlobalVariablesWithinANamespace1}, it looks complicated to me. The result of source~code~\ref{Code:GlobalVariablesWithinANamespace2} is the same, but in my personal opinion it's more readable and reminds me at e.g. C++.
\begin{lstlisting}[language=lua,float=htb,label=Code:GlobalVariablesWithinANamespace2,caption={Global variables within a namespace (2)}]
-- Create the "PublicVariables"-namespace
PublicVariables = {
	-- Create the variable "Health" within
	-- the "PublicVariables"-namespace
	Health	= 42,
	-- Create the variable "Gold" within
	-- the "PublicVariables"-namespace
	Gold	= 21,
}
\end{lstlisting}

When writing a script, it's probably a good idea to use such a \emph{public}-namespace to mark that those script content should be visible to e.g. an editor while other stuff is for internal purposes only. The Lua script language does also have the concept of local variables by writing the keyword \emph{local} in front of the variable declaration (see section~\ref{Script_GlobalVariables} for details). Local variables are invisible to the C++ host, so, they can also be used to realize script internal variables.

I hope you got the idea why using namespaces, even in scripts, is a good and handy thing.




\section{Global Variables}
\label{Script_GlobalVariables}
There's nothing fancy about global variables, so this section is quite compact. When using Lua, just assign somewhere within your script, as seen in source~code~\ref{Code:GlobalVariables}, a value to a name and et voila, you've created a new global variable\footnote{Lua is loosely typed, the type of a variable depends on the value assigned to it and can be changed every time by just assigning another value with another type to it}.
\begin{lstlisting}[language=lua,float=htb,label=Code:GlobalVariables,caption={Global variables}]
-- Create the global variable "Health"
Health = 42
-- Create the global variable "Gold"
Gold = 21
\end{lstlisting}

Lua does also have the concept of local variables by writing the keyword \emph{local} in front of the variable declaration as seen within source~code~\ref{Code:LocalVariables}.
\begin{lstlisting}[language=lua,float=htb,label=Code:LocalVariables,caption={Local variables}]
-- Create the global variable "Health"
Health = 42
-- Create the local variable "Gold"
local Gold = 21
\end{lstlisting}
From the C++ side, it's possible to interact with global variables while local variables can not be seen. So, while the C++ side can access your global variable \emph{Health}, it can't access your local variable \emph{Gold}.


\paragraph{Global Variables added by the C++ Host}
The C++ side is also able to create new, or to delete existing global variables. Whether or not this is in general useful is up to you. PixelLight components using scripts dynamically add a global variable named \emph{this} pointing back to the script calling \ac{RTTI} object. If you want to know more about the global \emph{this} variable, have a look into section~\ref{Script_RTTIObjects}.





\section{Global Functions}
\label{Script_GlobalFunctions}
As soon as your script grows over a hand full of source code lines, the need for functions arises. Within Lua, global functions look like as seen in source~code~\ref{Code:GlobalFunctions}.
\begin{lstlisting}[language=lua,float=htb,label=Code:GlobalFunctions,caption={Global functions}]
function MyFunction()
	-- ... do some fancy stuff...
end
\end{lstlisting}
The global function \emph{MyFunction} can then be called by writing \emph{MyFunction()}.

From the C++ side, it's possible to interact with global functions. Such global functions are usually used as script entry points and typical names for such global functions used as entry points are \emph{OnInit}, \emph{OnUpdate} and \emph{OnDeInit} which are called after as script has been loaded, when a script should perform an update step and shortly before a script is going to die. Please note that this was only an example of the naming convention used within PixelLight. Within the script system there are no such fixed build in names and you're free to choose your own entry point names. But in general, it makes the life easier to always use the same function names for the same purpose.


\paragraph{Global Functions added by the C++ Host}
Beside writing own global functions directly within the script, the C++ side is able to add global functions during runtime to expose stuff written in C++ to the script. This allows a script to communicate with it's C++ host. In general, global functions added by C++ should be kept to a bare minimum because they introduce script initialization cost and also trash the global script namespace, although there's namespace support to reduce the probability of name conflicts. A far more advanced, universal and powerful way of exposing C++ stuff to the script is by using \ac{RTTI} objects as discussed in section~\ref{Script_RTTIObjects}. Within the Lua script backend, the \ac{RTTI} object approach is not adding script initialization cost.


\paragraph{Functions used as Callbacks}
Just calling global function directly is not the only possible way to use them. Global functions can also be used as slots, meaning that when the signal they are connected with is emitted, the global function is called. Please note that this is also true for local functions, local functions can be used as slots as well. See section~\ref{Script_RTTIObjects_Slots} for details about slots and section~\ref{Script_RTTIObjects_Signals} for information about signals.

As seen within the following source~code~\ref{Code:VariableWithAFunctionAsValue}, within Lua it's also possible to use a function as a value of a variable.
\begin{lstlisting}[language=lua,float=htb,label=Code:VariableWithAFunctionAsValue,caption={Variable with a function as value}]
function MyFunction()
	-- ... do some fancy stuff...
end
myVariable = MyFunction
\end{lstlisting}
Therefore, a function can also be used as callback, name it \emph{slot} if you want. This way, a function pointer can be passed around within your script which can become really handy. Imagine that there's a movie script which accepts such a function pointer as parameter. As soon as the movie playback has been finished this given function is called. The script which is using the movie script can then react on this signal/event and change, for instance, into an interactive mode. Source~code~\ref{Code:FunctionAsCallback} summes this up.
\begin{lstlisting}[language=lua,float=htb,label=Code:FunctionAsCallback,caption={Function as callback}]
function MyCallbackFunction()
	-- ... do some fancy stuff...
end

function MyMovie(callbackFunction)
	-- ... do some fancy stuff...

	-- Movie playback has been finished
	if callbackFunction ~= nil then
		callbackFunction()
	end
end

-- Playback a movie
MyMovie(MyCallbackFunction)
\end{lstlisting}




\section{\ac{OOP}}
First at all, please note that this section has nothing to do with PixelLight \ac{RTTI} objects which are explained in section~\ref{Script_RTTIObjects}.

Let's directly jump into the topic. The \ac{OOP} Lua skeleton shown in source~code~\ref{Code:OOPSkeleton} has proven to be useful.
\begin{lstlisting}[language=lua,label=Code:OOPSkeleton,caption={\ac{OOP} Lua skeleton}]
--[-------------------------------------------------------]
--[ Includes                                              ]
--[-------------------------------------------------------]
-- Include another Lua script
require("FilenameOfAnotherLuaScript")


--[-------------------------------------------------------]
--[ Classes                                               ]
--[-------------------------------------------------------]
--@brief
--  My Lua class
MyClass = {


	--[-------------------------------------------------------]
	--[ Public definitions                                    ]
	--[-------------------------------------------------------]
	--@brief
	--   Interaction mode
	Mode = {
		WALK = 0,	-- Walk mode
		FREE = 1,	-- Free mode
	},


	--@brief
	--  The default constructor - In Lua a static method
	new = function()


		--[-------------------------------------------------------]
		--[ Private class attributes                              ]
		--[-------------------------------------------------------]
		-- A private class attribute -> Emulates the C++ "this"-pointer by using a Lua table
		local this	= {}
		-- The current interaction mode
		local _mode	= Interaction.Mode.WALK


		--[-------------------------------------------------------]
		--[ Private class methods                                 ]
		--[-------------------------------------------------------]
		--@brief
		--  Very secret stuff is happening in this private method
		local function MyPrivateMethod()
			-- ... do some fancy stuff...
		end


		--[-------------------------------------------------------]
		--[ Public class methods                                  ]
		--[-------------------------------------------------------]
		--@brief
		--  A public method
		function this.MyPublicMethod()
			-- ... do some fancy stuff...
		end


		--[-------------------------------------------------------]
		--[ Public class constructor implementation               ]
		--[-------------------------------------------------------]
		-- ... do some fancy stuff...


		-- Return the created class instance
		return this
	end


}
\end{lstlisting}
A new instance of this class can be created by writing \emph{myClassInstance = MyClass.new()} and methods can be called by writing \emph{myClassInstance.MyPublicMethod()}\footnote{Don't get confused by ''.'' and '':''. '':'' is a special Lua shortcut. For instance, instead of writing \emph{myClassInstance.MyPublicMethod(myClassInstance)} one can also write \emph{myClassInstance:MyPublicMethod()} to pass \emph{myClassInstance} as parameter into the method \emph{MyPublicMethod}.}.

Wait a moment, the class definition within source~code~\ref{Code:OOPSkeleton} just looks like a namespace as described within section~\ref{Script_Namespaces}... and hey, Lua doesn't have \ac{OOP} support at all!

That's true, Lua has no build in \ac{OOP} support, but that's no reason to don't use \ac{OOP}. It's possible to realize \ac{OOP} by using, yes your guess was right, tables. Even information hiding is possible.

At this point you may possibly ask yourself: \begin{quote}''Why use \ac{OOP} at all within a script? Shouldn't make a script language the life of a programmer much easier? Why can't we just use variables and functions and be happy?''\end{quote} The answer is short: No one is forcing you to use \ac{OOP}. If you're uncomfortable or unfamiliar with \ac{OOP} and don't want to get into the topic, then don't. If your scripted application gets bigger and you want a powerful tool to write well structured, reusable script code without a totally flooded global namespace due to thousands of global variables and functions, use the mentioned \ac{OOP} approach. Personally, I wouldn't want to work without \ac{OOP} because I'am a lazy person.




\section{\ac{RTTI} Objects}
\label{Script_RTTIObjects}
\ac{RTTI} objects is, were the scripting gets really interesting because it's possible to directly access and manipulate PixelLight \ac{RTTI} C++ objects by using a script. Within the Lua script backend, \ac{RTTI} objects are realized by using the \emph{user data} feature of Lua together with Lua's metatables. Whenever Lua is not able to resolve for instance a function call, this request is redirected into the script backend implementation which is then accessing the \ac{RTTI} system of PixelLight.

\paragraph{this}
PixelLight components using scripts dynamically add a global variable named \emph{this} pointing back to the script calling \ac{RTTI} object. Of course, the type of the \ac{RTTI} object depends on the use-case. For example when running a scripted stand-alone application by using \emph{PLViewer}, \emph{this} points to an instance of the \emph{PLEngine::ScriptApplication} \ac{RTTI} class. Another example would be the script scene node modifier called \emph{PLScriptBindings::SNMScript} where \emph{this} points to an instance of \emph{PLScriptBindings::SNMScript}. All \ac{RTTI} objects are derived from the base class \emph{PLCore::Object} and therefore share a common feature set. To check the \ac{RTTI} object type, just write within a Lua script \emph{myRTTIObject:IsInstanceOf("PLEngine::ScriptApplication")} in order to figure out whether the \ac{RTTI} object you currently dealing with is an instance of \emph{PLEngine::ScriptApplication}. In case you've wondered why it's called \emph{this} and not \emph{self} as it's commonly named in Lua scripts: The answer is, that PixelLight is using a generic script language independent script interface. Components like \emph{PLEngine::ScriptApplication} or \emph{PLScriptBindings::SNMScript} don't use e.g. Lua directly, they're only using the generic script language independent script interface. Not every script language is using \emph{self} to identify \emph{this instance}. PixelLight itself is written in C++, so the name \emph{this} was obviously a better choice then \emph{self} because this way, there are not to many different names describing one and the same concept.


\subsection{Properties}
Within PixelLight, \ac{RTTI} class properties are constant \ac{RTTI} class information which are defined within C++ as seen in source~code~\ref{Code:RTTIClassPropertiesDefinition} (see PLCore documentation for details).
\begin{lstlisting}[label=Code:RTTIClassPropertiesDefinition,caption={Defining a new \ac{RTTI} class with properties (C++)}]
#include <PLCore/Base/Object.h>
class MyClass : public PLCore::Object {
	pl_class(MY_RTTI_EXPORT, MyClass, "", PLCore::Object, "Description of my RTTI class")
		// Properties
		pl_properties
			pl_property("MyFirstProperty",  "FirstValue")
			pl_property("MySecondProperty", "SecondValue")
		pl_properties_end
		// Constructors
		pl_constructor_0(MyConstructor, "Default constructor", "")
	pl_class_end
};
\end{lstlisting}

Within a script language, these properties can be accessed by writing \emph{<property value> = <RTTI object>.<property name>}. For example, if your \ac{RTTI} object instance is named \emph{myRTTIObject} and you want to get the value of the property with the name \emph{MyFirstProperty} then just write \begin{quote}value = myRTTIObject.MyFirstProperty\end{quote}


\subsection{Attributes}
Within PixelLight, \ac{RTTI} class attributes are member variables of a class which are defined within C++ as seen in source~code~\ref{Code:RTTIClassWithAttributes} (see PLCore documentation for details).
\begin{lstlisting}[label=Code:RTTIClassWithAttributes,caption={Defining a new \ac{RTTI} class with attributes (C++)}]
// Class definition of MyClass
#include <PLCore/Base/Object.h>
class MyClass : public PLCore::Object {

	// RTTI interface
	pl_class(MY_RTTI_EXPORT, MyClass, "", PLCore::Object, "Description")
		// Attributes
		pl_attribute(MyInt, int, 10, ReadWrite, DirectValue, "Simple integer", "")
		pl_attribute(MyFloat, float, 3.1415f, ReadOnly, DirectValue, "PI", "")
		// Constructors
		pl_constructor_0(MyConstructor, "Default constructor", "")
	pl_class_end

	// Default constructor
	public:
		MyClass() : MyInt(this), MyFloat(this) {}

};

// MyClass RTTI implementation (not done within headers)
pl_implement_class(MyClass)
\end{lstlisting}

Within a script language, these attributes can be accessed by writing \emph{<attribute value> = <RTTI object>.<attribute name>} to read a value and by writing \emph{<RTTI object>.<attribute name> = <attribute value>} to write a value. For example, if your \ac{RTTI} object instance is named \emph{myRTTIObject} and you want to get the value of the attribute with the name \emph{MyInt} then just write \begin{quote}value = myRTTIObject.MyInt\end{quote}


\subsection{Methods}
Within PixelLight, \ac{RTTI} class methods are member functions of a class which are defined within C++ as seen in source~code~\ref{Code:RTTIMethods} (see PLCore documentation for details).
\begin{lstlisting}[label=Code:RTTIMethods,caption={Definition of \ac{RTTI} class methods with parameters (C++)}]
// Class definition of MyClass
#include <PLCore/Base/Object.h>
class MyClass : public PLCore::Object {

	// RTTI interface
	pl_class(MY_RTTI_EXPORT, MyClass, "", PLCore::Object, "Description of my RTTI class")
		// Constructors
		pl_constructor_0(MyConstructor, "Default constructor", "")
		// Methods
		pl_method_0(MyMethod, pl_ret_type(void), "My method", "")
		pl_method_1(MyMethodParameter, pl_ret_type(void), int, "My method with given parameter", "")
		pl_method_2(MyMethodParameters, bool, int, float, "My method with given parameters and return value", "")
	pl_class_end

	public:
		MyClass() {}
		void MyMethod() {}
		void MyMethodParameter(int nFirst) {}
		bool MyMethodParameters(int nFirst, float fSecond) { return true; }

};

// MyClass RTTI implementation
pl_implement_class(MyClass)
\end{lstlisting}

Within a script language, these methods can be accessed by writing \emph{<result> = <RTTI object>:<method name>(<parameters>)}. For example, if your \ac{RTTI} object instance is named \emph{myRTTIObject} and you want to call the method with the name \emph{MyMethodParameters} then just write \begin{quote}result = myRTTIObject:MyMethodParameters(42, 21.21)\end{quote}

Please note that the result of such an \ac{RTTI} object method call can be an \ac{RTTI} object. You're able to call methods of this returned \ac{RTTI} object as well... and from \ac{RTTI} objects returned by this call and so on.


\subsection{Slots}
\label{Script_RTTIObjects_Slots}
Within PixelLight, \ac{RTTI} class slots are event handlers within a class which are defined within C++ as seen in source~code~\ref{Code:RTTISlots} (see PLCore documentation for details).
\begin{lstlisting}[label=Code:RTTISlots,caption={Definition of \ac{RTTI} class slots with parameters (C++)}]
// Class definition of MyClass
#include <PLCore/Base/Object.h>
class MyClass : public PLCore::Object {

	// RTTI interface
	pl_class(MY_RTTI_EXPORT, MyClass, "", PLCore::Object, "Description of my RTTI class")
		// Constructors
		pl_constructor_0(MyConstructor, "Default constructor", "")
		// Slots
		pl_slot_0(OnMyEvent, "My slot", "")
		pl_slot_1(OnMyEventParameter, int, "My slot with given parameter", "")
	pl_class_end

	// Public methods
	public:
		MyClass() : SlotOnMyEvent(this), SlotOnMyEventParameter(this) {}
		void OnMyEvent() {}
		void OnMyEventParameter(int nValue) {}

};

// MyClass RTTI implementation
pl_implement_class(MyClass)
\end{lstlisting}

Within a script language, these slots can be connect to and disconnect from a \ac{RTTI} class signal. Have a look at the following section~\ref{Script_RTTIObjects_Signals} to see how this can be done.


\subsection{Signals}
\label{Script_RTTIObjects_Signals}
Within PixelLight, \ac{RTTI} class signals are events within a class which are defined within C++ as seen in source~code~\ref{Code:RTTISignalDefinition} (see PLCore documentation for details).
\begin{lstlisting}[label=Code:RTTISignalDefinition,caption={Defining a \ac{RTTI} class with signals (C++)}]
// Class definition of MyClass
#include <PLCore/Base/Object.h>
class MyClass : public PLCore::Object {

	// RTTI interface
	pl_class(MY_RTTI_EXPORT, MyClass, "", PLCore::Object, "Description of my RTTI class")
		// Constructors
		pl_constructor_0(MyConstructor, "Default constructor", "")
		// Signals
		pl_signal_0(MySignal, "My signal", "")
		pl_signal_1(MySignalParameter, int, "My signal with given parameter", "")
	pl_class_end

	// Public methods
	public:
		MyClass() {}

};

// MyClass RTTI implementation
pl_implement_class(MyClass)
\end{lstlisting}

Within a script language, these signals can be accessed by writing \emph{<RTTI object>:<signal name>(<parameters>)} to emit the signal. For example, if your \ac{RTTI} object instance is named \emph{myRTTIObject} and you want to emit the signal with the name \emph{MySignalParameter} then just write \begin{quote}myRTTIObject:MySignalParameter(42)\end{quote}

Within a script, signals can be connected to an \ac{RTTI} object slot (see section~\ref{Script_RTTIObjects_Slots} for details about \ac{RTTI} slots) or to a global function (section~\ref{Script_GlobalFunctions} for information about global functions) with then is used as a script slot. There are two fixed build in script methods for signals:
\begin{itemize}
\item{\emph{Connect} is used to connect a slot, which is provided as parameter, to the signal}
\item{\emph{Disconnect} is used to disconnect a slot, which is provided as parameter, from the signal}
\end{itemize}

For example, if your \ac{RTTI} object instance is named \emph{myRTTIObject} and you want to connect the \ac{RTTI} class slot with the name \emph{MySlot} of the same \ac{RTTI} object to a signal with the name \emph{MySignalParameter}, then just write \begin{quote}myRTTIObject.MySignalParameter.Connect(myRTTIObject.MySlot)\end{quote}

As mentioned before, global as well as local script functions can also be used as slots. If the \ac{RTTI} class signal is emitted, the global script function will be called. Source~code~\ref{Code:FunctionAsSlot} shows how this can be used in practice.
\begin{lstlisting}[language=lua,float=htb,label=Code:FunctionAsSlot,caption={Function as slot}]
-- Global script function which is used as slot
function MySlotFunction()
	-- ... do some fancy stuff...
end

-- Use the script function "MySlotFunction" as slot and
-- connect it with the RTTI signal of our RTTI class instance
myRTTIObject.MySignalParameter.Connect(MySlotFunction)

-- Emit the RTTI signal
myRTTIObject:MySignalParameter(42)
\end{lstlisting}


\subsection{Creating a \ac{RTTI} Class Instance via Script}
It's possible to create a \ac{RTTI} class instance via script. This instance is then managed by the script and will automatically be destroyed when it's no longer referenced. For the purpose of creating a \ac{RTTI} class instance via script we're providing the \emph{PL.ClassManager.CreateByConstructor()}\footnote{See chapter~\ref{ScriptBindingsPlugin} for more information about available script bindings} function which can be used as seen within source~code~\ref{Code:CreatingARTTIClassInstanceViaScript}.
\begin{lstlisting}[language=lua,float=htb,label=Code:CreatingARTTIClassInstanceViaScript,caption={Creating a RTTI Class Instance via Script}]
cppCamcorder = PL.ClassManager.CreateByConstructor("PLEngine::Camcorder", "ParameterConstructor", "Param0=\"" .. tostring(PL.GetApplication()) .. "\"")
cppCamcorder:StartCamcorderPlayback("MyAwesomeMovie")
\end{lstlisting}
This example creates an instance of the \emph{PLEngine::Camcorder} class, the constructor is provided with a pointer to the C++ application class requested via the script binding function \emph{PL.GetApplication()}. The created \ac{RTTI} class instance can be used as usual as seen in the presented example. You don't need to care about the fact that this instance was created by the script. By using this approach, it's possible to create more complex or performance critical loose components within C++, and then instancing and using it easily from inside a script.
