%----- Chapter: Introduction ---------------------
\chapter{Introduction}


\paragraph{Motivation}
The PixelLight core component provides more advanced features that are usually quite useful for all types of projects. As the project name \emph{PLCore} implies, it's the foundation of the PixelLight framework which is used for the development of complete applications. RTTI, plugins, the application framework - this all is part of the \emph{PLCore} project.


\paragraph{Overview}
The heart of PLCore is the RTTI which is completely basing on C++ templates and additionally offers some macros to enhance the readability and usability. By using C++ templates, the system is type safe. The RTTI is using several fundamental components like variables, functions, events and event handler. This components, the RTTI is composed of, can be subdivided into three levels as summarized within table~\ref{Table:FundamentalRTTIComponents}.
\begin{table}[htb]
	\centering
	\begin{tabular}{|l||l|l|l|}
		\hline
		Components & 1. Virtual Base Class & 2. Typed Variant & 3. Inside Object\\
		\hline
		\hline
		Variable & DynVar & Var\textless type\textgreater & Attribute\textless type\textgreater\\
		\hline
		Function & DynFunc & Func\textless type\textgreater & Method\textless type\textgreater\\
		\hline
		Event & DynEvent & Event\textless type\textgreater & Signal\textless type\textgreater\\
		\hline
		EventHandler & DynEventHandler & EventHandler\textless type\textgreater & Slot\textless type\textgreater\\
		\hline
	\end{tabular} 
	\caption{Fundamental RTTI components and their levels}
	\label{Table:FundamentalRTTIComponents}
\end{table}


\paragraph{'this' Used In Constructor Initializer List}
When you try to compile examples shown within this document, you will probably get a compiler warning like \begin{quote}warning C4355: 'this' : used in base member initializer list\end{quote}. This warning occurs because within the initializer-list, the current object denoted by \emph{this} may not be fully initialized, yet. In here, you can just ignore this warning because internally the given pointer is just stored but not yet accessed - so, there is no immediate danger in using \emph{this}. Source~code~\ref{Code:DisableThisInitializerListWarning} shows how this warning can be disabled within \emph{Microsoft Visual Studio 2008}.
\begin{lstlisting}[float=htb,label=Code:DisableThisInitializerListWarning,caption={Disable warning when using \emph{this} within the initializer list}]
// Disable warning
PL_WARNING_PUSH
PL_WARNING_DISABLE(4355) // "'this' : used in base member initializer list"

	// Default constructor
	MyClass::MyClass() :
		MyInt(this),
		MyFloat(this)
	{
	}

// Reset to previous warning settings
PL_WARNING_POP
\end{lstlisting}
While doing this for each case in a controlled way would be the clean way - it would also mean that, for each class, there would be more typing work and the code wouldn't look that compact any more. Therefore, we decided to disable this one warning by default within \emph{PLCoreWindows.h} - usually we don't do that because warnings are in general there for a good and helpful reason, but in this case the alternative wouldn't be better.




\section{External Dependences}
PLCore only depends on the \textbf{PLGeneral} library and doesn't use any external third party libraries itself.
